{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory

### 2024-03-18 paper/main.tex
alphabetize keywords

~\\
{\it Keywords:} Evolution of cooperation; direct reciprocity; repeated prisoner's dilemma; social learning; evolutionary dynamics

@@

Most models of direct reciprocity abstract from these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,hilbe:PNAS:2017,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023}. 
They just assume individuals can easily copy the strategies of others. 
Similarly, they just assume that updating decisions are based on the strategies' average (or expected) payoffs, which are based on all rounds and all interactions. 
These assumptions create a curious inconsistency in how models represent an individual's cognitive abilities. 
On the one hand, when playing the game, individuals are often assumed to have restricted memory. 
Respective studies typically assume that individuals make their decisions each round based on the outcome of the last round only~\citep[with only a few exceptions, see Refs.][]{Hauert1997,van-veelen:PNAS:2012,Stewart2016,Li:NatCS:2022,Murase:PLoSCompBio:2023a}. 
Yet when learning new strategies, individuals are assumed to remember (or compute) each others' precise average payoff across many rounds and many interaction partners. 
Herein, we wish to explore whether this latter assumption is actually necessary for the evolution of reciprocity through social imitation. 
We ask whether individuals can learn to adopt reciprocal strategies even when learning is based on payoff information from a limited number of rounds.

@@

\noindent
{\bf Description of the game dynamics.} We consider a well-mixed population consisting of $N$~players.
Players are randomly matched in pairs to participate in a repeated donation game~\citep{sigmund2010calculus} with their respective co-player.
Each round, they can either cooperate (\(C\)) or defect (\(D\)). 
A cooperating player provides a benefit~\(b\) to the other player at their own cost~\(c\), with \(0 \!<\! c \!<\! b\). 
A defecting player provides no benefit and pays no cost. 
Thus, the players' payoffs in a single round are given by the matrix
\begin{align}
\bordermatrix{%
	& C & D \cr
C &\ b-c &\ -c\  \cr
D &\ b &\ 0\ \cr
} .
\end{align}
In particular, payoffs take the form of a prisoner's dilemma:
Mutual cooperation yields a better payoff than mutual defection ($b\!-\!c\!>\!0$), but each player individually prefers to defect independent of the co-player's action ($b\!>\!b\!-\!c$ and $0\!>\!-c$). 
To incorporate that individuals interact repeatedly, we assume that after each round, there is a constant continuation probability $\delta$ of interacting for another round. 
For $\delta\!=\!0$, we recover the case of a conventional (one-shot) prisoner's dilemma. 
Here, mutual defection is the only equilibrium. 
As $\delta$ increases, the game turns into a repeated game. Here, additional equilibria emerge, with some of them allowing for full cooperation~\citep{friedman:RES:1971,Akin:chapter:2016,hilbe:GEB:2015,stewart:pnas:2014}.

@@

Both in the case of perfect and limited memory, we iterate the elementary strategy updating step described above for many time steps. 
This gives rise to a stochastic process that describes which strategies players adopt over time. 
We explore the dynamics of this process mathematically and with computer simulations.
For the results presented in the following, we assume that mutations are rare (\(\mu\!\rightarrow\! 0\)). 
This assumption is fairly common in evolutionary game theory, because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}, and because the results can be interpreted more easily.  
However, in Section~3 of the \esm{} we show that our main results continue to hold for strictly positive mutation rates.

@@

In the case of perfect payoff memory, it is straightforward to characterize when cooperation is stochastically stable. 
Here, we simply need to compute the players' expected payoffs. 
Because the population mostly consists of residents, and because residents mutually cooperate with each other, their expected payoff is $\pi_\gtft = b\!-\!c$. 
On the other hand, the defecting mutant only interacts with residents. 
Given the residents' strategy, the mutant receives a benefit in the first round, and in every subsequent round with probability $q$. 
As a result, the mutant's expected payoff is $\pi_\alld \!=\! (1\!-\!\delta\!+\!\delta q)b$. 
For perfect payoff memory, the requirement for cooperation to be stochastically stable reduces to the condition $\pi_\gtft > \pi_\alld$. 
This yields
\begin{equation} \label{Eq:PerfectMemory}
q < 1\!-\!\frac{c}{\delta  b}.
\end{equation}
In particular, we recover the previous observation that $q\!=\!1-c/(\delta b)$ is the maximum generosity that cooperators should have~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}. 
Because $q\!\ge\!0$, we also conclude that cooperation can only be stable if $\delta \!>\! c/b$.
Again, this condition for the feasibility of direct reciprocity is the condition found in the literature~\citep{nowak:Science:2006}.

@@

\noindent
{\bf Evolutionary dynamics of reciprocity.}
To explore whether the previous static observations describe the dynamics of evolving populations, we turn to simulations.
We have run separate simulations for perfect and limited payoff memory. 
In each case, we consider both a low and a high benefit of cooperation ($b/c\!=\!3$ and $b/c\!=\!10$, respectively).  
For each simulation, we record which strategies ($y,p,q$) the players adopt over time.
\FigBaseResults{} depicts the conditional cooperation probabilities $p$ and $q$ (we omit the opening move \(y\) because we use a discount factor~\(\delta\) close to one, such that first-round behavior is largely irrelevant). 
In all simulations, we find that the players' strategies cluster in two regions of the strategy space. 
The first region corresponds to a neighborhood of \alld{} with $(p,q)\!\approx\!(0,0)$.
The second region corresponds to a thin strip of cooperative strategies with $(p,q)\!\approx\!(1,q)$. 
Within this strip, we observe that most strategies satisfy the constraints on $q$ imposed by the inequalities~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}. 
That is, with perfect memory, most evolving strategies have $q\!<\!1\!-\!c/b$, whereas with limited payoff memory, most strategies have $q\!<\!1/2$. 
In particular, for limited payoff memory, changes in the benefit parameter have no effect on the qualitative distribution of strategies.

@@

In each case, the evolutionary dynamics follow a similar cyclic pattern~\citep[as described in Refs.][]{imhof2010stochastic, Nowak1992tit}:
Resident populations of defectors are most likely invaded by strategies close to \tft. 
Once the population adopts conditionally cooperative strategies  $(1,1,q)$, neutral drift may introduce larger values of generosity~$q$. 
If the resident's generosity~$q$ violates the conditions~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}, defectors can re-invade, and the cycle starts again. 
The relative time spent near \alld{} and near the strip of conditionally cooperative strategies depends on the considered memory setting (\FigInvasionAnalysis, depicting the case of high benefits). 
For perfect memory, we find that \alld{} is replaced relatively quickly by more cooperative strategies. 
Here, it takes on average 159 invading mutants until \alld{} is successfully replaced. 
In contrast, for limited memory, \alld{} is more robust, resisting on average 798 mutant strategies. 
This picture reverses when we consider an initial population that adopts \gtft. 
Such populations are much more robust under perfect memory than they are under limited memory. 
Overall, we find that the impact of memory on the population's average cooperation rate is substantial. 
For perfect memory, this rate is 52\% for low benefits, and 98\% for high benefits. 
For limited payoff memory, the evolving cooperation rates are smaller but still strictly positive, with 37\% cooperation for low benefits and 51\% cooperation for high benefits (\FigBaseResults).

@@

To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in~\FigDependenceParameters.
According to \FigDependenceParameters{\it a}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
This again reflects our earlier observation that the feasibility of cooperation in this scenario is largely independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}. 
With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yield similar cooperation rates for weak selection (\(\beta \!\ll\! 1\)). This is not a coincidence, since it is known that, under weak selection, stochastic payoffs can be replaced by their (deterministic) expectations without altering the evolutionary dynamics \citep{mcavoy:NHB:2020}--and perfect payoff memory corresponds to the expected value of the payoffs in the limited payoff memory model, due to the law of large numbers.
Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\

@@

\noindent
{\bf Beyond reactive strategies.} While the results presented in the main text focus on reactive strategies, the patterns we observe do not seem to depend on the considered strategy space. 
To illustrate this point in more detail, in the \esm{} we consider the dynamics among memory-1 strategies. 
Here, players take into account both their co-player's and their own last move, see Refs.~\citep{nowak:Nature:1993,imhof:JTB:2007}. 
Also in that case, we observe that perfect memory leads to systematically higher cooperation rates (Supplementary Figures~5,6).
Again, this advantage of perfect memory is particularly pronounced for strong selection, or when there is a high benefit of cooperation (\FigMemoryOneParameters).\\

@@

\noindent 
{\bf The effect of increasing individual payoff memory.}
So far, we have taken a rather extreme interpretation of limited payoff memory. 
In the respective scenario, we assumed that individuals update their strategies based on their experience in a single round of the prisoner's dilemma, against a single co-player. 
The limited payoff memory framework can be expanded in various ways. 
In particular, individuals may recall a larger number of rounds, they may recall their interactions with several co-players, or both. 
To gain further insights on the impact of payoff memory, we explore four additional scenarios. 
In the first scenario, players recall the payoffs they obtained in the last two rounds against a single co-player. 
In the second scenario, players recall their last-round payoffs against two co-players. 
In the third scenario, they recall the two last rounds against two co-players. 
Finally, in the last scenario, players update based on the average payoff they
receive over all rounds with a single co-player (further extensions are possible, but we do not explore them here).

@@

We complement these analytical results with additional simulations, see \FigHigherMemory.
We observe that a minimal increase in the players' payoff memory (compared to the baseline case with a single round recalled) can promote cooperation considerably. 
Specifically, in all four scenarios with extended memory, we see similar cooperation rates, and they approach the rates observed under perfect memory. 
These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest. 
Already remembering a few interactions, either with the same co-player or across different co-players, may provide players with enough information to adopt reciprocal strategies.

%% RESULTS: Snowdrift Game
{\bf Beyond the donation game.}
So far, we have discussed the donation game, which is the most common
example of social dilemmas. There are several other examples, which include the
snowdrift game~\cite{doebeli:Ecology:2005}, the stag hunt
game~\cite{pacheco:ProcB:2009}, and the public goods
game~\cite{pinheiro:PLoSCompBio:2014}. However, the pattern we observe is more
general and still holds for other social dilemmas. To demonstrate this, we
consider the snowdrift game. In the snowdrift game, individuals can again
cooperate or defect. However, now if both cooperate, they have the benefit \(b\)
of getting home while sharing the labor \(c\). Thus for mutual cooperation, they
receive \(b - \frac{c}{2}\). However, if one defects, they get \(b\) while
the cooperator gets \(b - c\). By contrast to the donation game, now
the best action depends on the co-player: to defect if the other cooperates, but
to cooperate if the other defects.

@@

Apart from the analytical analysis, we evaluate our results with simulations,
and we do this for all the updating payoff cases considered in this manuscript.
What we find is that the qualitative results still hold for the snowdrift game.
Also, in that case, we observe that perfect memory leads to systematically
higher cooperation rates (Supplementary Figures~8, 9). However, now less memory
is needed to approximate the classical limit.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=.75\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
    \caption{{\bf Evolution of direct reciprocity for different parameter values.} 
    To explore the robustness of our results, we have run simulations for different benefit values (left panels, {\it a}), and for different selection strengths (right panels, {\it b}). 
    In each case, we record the resulting average cooperation rate over the entire simulation (upper panels). 
    In addition we record the individuals' average generosity. 
    Here, we only take into account those residents with $p\! \approx\! 1$ and we compute
    the average of their cooperation probability~$q$. 
    These simulations suggest that perfect payoff memory consistently leads to more cooperation and more generosity. 
    Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
    Each point represents a single simulation run. Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter combination. Each point is a sinlge simualtion run.}
    \label{fig:cooperation_rate_over_benefit_and_beta}
\end{figure}

>>>>

~\\
{\it Keywords:} Evolution of cooperation; evolutionary dynamics; direct reciprocity; repeated prisoner's dilemma; social learning

@@

Most models of direct reciprocity do not address these difficulties~\citep{brauchli:JTB:1999,brandt:JTB:2006,ohtsuki:JTB:2007b,szolnoki:pre:2009b,imhof2010stochastic,van-segbroeck:prl:2012,grujic:jtb:2012,Martinez2012,stewart:pnas:2013,pinheiro:PLoSCB:2014,stewart:games:2015,Baek2016,hilbe:PNAS:2017,McAvoy:ProcA:2019,glynatsi:SCR:2020,Schmid:PlosCB:2022,Murase:SciRep:2022,Cooney:BMB:2022,Chen:PNASNexus:2023} and, instead, assume individuals can easily copy the strategies of others. 
Similarly, they assume that updating decisions are based on the strategies' average (or expected) payoffs, which are based on all rounds and all interactions. 
These assumptions create a curious inconsistency in how models represent an individual's cognitive abilities. 
On the one hand, when playing the game, individuals are often assumed to have restricted memory. 
Respective studies typically assume that individuals make their decisions each round based on the outcome of the last round only~\citep[with only a few exceptions, see Refs.][]{Hauert1997,van-veelen:PNAS:2012,Stewart2016,Li:NatCS:2022,Murase:PLoSCompBio:2023a}. 
Yet when learning new strategies, individuals are assumed to remember (or compute) each others' precise average payoff across many rounds and many interaction partners. 
Herein, we wish to explore whether this latter assumption is actually necessary for the evolution of reciprocity through social imitation. 
We ask whether individuals can learn to adopt reciprocal strategies even when learning is based on payoff information from a limited number of rounds.

@@

\noindent
{\bf Description of the game dynamics.} We consider a well-mixed population consisting of $N$~players.
Players are randomly matched in pairs to participate in a repeated donation game~\citep{sigmund2010calculus} with their respective co-player.
Each round, they can either cooperate (\(C\)) or defect (\(D\)). 
A cooperating player provides a benefit~\(b\) to the other player at their own cost~\(c\), with \(0 \!<\! c \!<\! b\). 
A defecting player provides no benefit and pays no cost. 
Thus, the players' payoffs in a single round are given by the matrix
\begin{align}
\bordermatrix{%
	& C & D \cr
C &\ b-c &\ -c\  \cr
D &\ b &\ 0\ \cr
} .
\end{align}
In particular, payoffs take the form of a prisoner's dilemma:
Mutual cooperation yields a better payoff than mutual defection ($b\!-\!c\!>\!0$), but each player individually prefers to defect, independent of the co-player's action ($b\!>\!b\!-\!c$ and $0\!>\!-c$). 
To incorporate repetition, we assume that after each round, there is a constant continuation probability $\delta$ of interacting for another round. 
For $\delta\!=\!0$, we recover the case of a conventional (one-shot) prisoner's dilemma. 
Here, mutual defection is the only equilibrium. 
As $\delta$ increases, the game turns into a repeated game. Here, additional equilibria emerge, with some of them allowing for full cooperation~\citep{friedman:RES:1971,Akin:chapter:2016,hilbe:GEB:2015,stewart:pnas:2014}.

@@

Both in the case of perfect and limited memory, we iterate the elementary strategy update step described above many times. 
This gives rise to a stochastic process that describes which strategies players adopt over time. 
We explore the dynamics of this process mathematically and with numerical simulations.
For the results presented in the following, we assume that mutations are rare (\(\mu\!\rightarrow\! 0\)). 
This assumption is fairly common in evolutionary game theory, both because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015} and because the results can be interpreted more easily.  
However, in Section~3 of the \esm{} we show that our main results continue to hold for strictly positive mutation rates.

@@

In the case of perfect payoff memory, it is straightforward to characterize when cooperation is stochastically stable. 
Here, we simply need to compute the players' expected payoffs. 
Because the population mostly consists of residents, and because residents mutually cooperate with each other, their expected payoff is $\pi_\gtft = b\!-\!c$. 
On the other hand, the defecting mutant only interacts with residents. 
Given the residents' strategy, the mutant receives a benefit in the first round, and in every subsequent round with probability $q$. 
As a result, the mutant's expected payoff is $\pi_\alld \!=\! (1\!-\!\delta\!+\!\delta q)b$. 
For perfect payoff memory, the requirement for cooperation to be stochastically stable reduces to the condition $\pi_\gtft > \pi_\alld$, which yields
\begin{equation} \label{Eq:PerfectMemory}
q < 1\!-\!\frac{c}{\delta  b}.
\end{equation}
In particular, we recover the previous observation that $q\!=\!1-c/(\delta b)$ is the maximum generosity that cooperators should have~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}. 
Because $q\!\ge\!0$, we also conclude that cooperation can be stable only if $\delta \!>\! c/b$.
Again, this condition for the feasibility of direct reciprocity already exists in the literature~\citep{nowak:Science:2006}.

@@

\noindent
{\bf Evolutionary dynamics of reciprocity.}
To explore whether the previous observations describe the dynamics of evolving populations, we turn to simulations.
We have run separate simulations for perfect and limited payoff memory. 
In each case, we consider both a low and a high benefit of cooperation ($b/c\!=\!3$ and $b/c\!=\!10$, respectively).  
For each simulation, we record which strategies ($y,p,q$) the players adopt over time.
\FigBaseResults{} depicts the conditional cooperation probabilities $p$ and $q$ (we omit the opening move \(y\) because we use a discount factor~\(\delta\) close to one, such that first-round behavior is largely irrelevant). 
In all simulations, we find that the players' strategies cluster in two regions of the strategy space. 
The first region corresponds to a neighborhood of \alld{} with $(p,q)\!\approx\!(0,0)$.
The second region corresponds to a thin strip of cooperative strategies with $(p,q)\!\approx\!(1,q)$. 
Within this strip, we observe that most strategies satisfy the constraints on $q$ imposed by the inequalities~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}. 
That is, with perfect memory, most evolving strategies have $q\!<\!1\!-\!c/b$, whereas with limited payoff memory, most strategies have $q\!<\!1/2$. 
In particular, for limited payoff memory, changes in the benefit parameter have no effect on the qualitative distribution of strategies.

@@

In each case, the evolutionary dynamics follow a similar cyclic pattern~\citep[as described in Refs.][]{imhof2010stochastic, Nowak1992tit}:
Resident populations of defectors are most likely invaded by strategies close to \tft. 
Once the population adopts conditionally cooperative strategies  $(1,1,q)$, neutral drift may introduce larger values of generosity~$q$. 
If the resident's generosity~$q$ violates the conditions~\eqref{Eq:PerfectMemory} and~\eqref{Eq:LimitedMemory}, defectors can re-invade, and the cycle starts again. 
The relative time spent near \alld{} and near the strip of conditionally cooperative strategies depends on the memory (\FigInvasionAnalysis, depicting the case of high benefits). 
For perfect memory, we find that \alld{} is replaced relatively quickly by more cooperative strategies. 
Here, it takes on average 159 invading mutants until \alld{} is successfully replaced. 
In contrast, for limited memory, \alld{} is more robust, resisting on average 798 mutant strategies. 
This picture reverses when we consider an initial population that adopts \gtft. 
Such populations are much more robust under perfect memory than they are under limited memory. 
Overall, we find that the impact of memory on the population's average cooperation rate is substantial. 
For perfect memory, this rate is 52\% for low benefits, and 98\% for high benefits. 
For limited payoff memory, the evolving cooperation rates are smaller but still strictly positive, with 37\% cooperation for low benefits and 51\% cooperation for high benefits (\FigBaseResults).

@@

To further investigate the influence of different parameters, we have systematically varied the benefit~$b$ and the selection strength~$\beta$ in~\FigDependenceParameters.
According to \FigDependenceParameters{\it a}, perfect memory consistently results in a higher cooperation rate, and this relative advantage further increases with an increasing benefit~$b$. 
Interestingly, for limited payoff memory, the cooperation rate remains stable at approximately 50\% once \(b \!\ge\! 5\).
This again reflects our earlier observation that the feasibility of cooperation in this scenario is largely independent of the exact values of $b$ and $c$, as described by Eq.~\eqref{Eq:LimitedMemory}. 
With respect to the effect of different selection strengths, \FigDependenceParameters{\it b} suggests that both perfect and limited payoff memory yield similar cooperation rates for weak selection (\(\beta \!\ll\! 1\)). This is not a coincidence: it is known that, under weak selection, stochastic payoffs can be replaced by their (deterministic) expectations without altering the evolutionary dynamics \citep{mcavoy:NHB:2020}--and perfect payoff memory corresponds to the expected value of the payoffs in the limited payoff memory model, due to the law of large numbers.
Beyond weak selection, increasing selection has a positive effect under perfect payoff memory, but a negative effect under limited payoff memory.\\

@@

\noindent
{\bf Beyond reactive strategies.} While the results presented in the main text focus on reactive strategies, the patterns we observe do not appear to be sensitive to the strategy space. 
To illustrate this point in more detail, in the \esm{} we consider the dynamics among memory-1 strategies. 
Here, players take into account both their co-player's and their own last move, see Refs.~\citep{nowak:Nature:1993,imhof:JTB:2007}. 
In that case, we also observe that perfect memory leads to systematically higher cooperation rates (Supplementary Figures~5,6).
Again, this advantage of perfect memory is particularly pronounced for strong selection, or when there is a high benefit of cooperation (\FigMemoryOneParameters).\\

@@

\noindent 
{\bf The effect of increasing individual payoff memory.}
So far, we have taken a rather extreme interpretation of limited payoff memory. 
In the respective scenario, we assumed that individuals update their strategies based on their experience in a single round of the prisoner's dilemma, against a single co-player. 
The limited payoff memory framework can be expanded in various ways. 
In particular, individuals may recall a larger number of rounds, they may recall their interactions with several co-players, or both. 
To gain further insights on the impact of payoff memory, we explore four additional scenarios. 
In the first scenario, players recall the payoffs they obtained in the last two rounds against a single co-player. 
In the second scenario, players recall their last-round payoffs against two co-players. 
In the third scenario, they recall the two last rounds against two co-players. 
Finally, in the last scenario, players update based on the average payoff they
receive over all rounds with a single co-player. (Further extensions are possible, but we do not explore them here.)

@@

We complement these analytical results with additional simulations; see \FigHigherMemory.
We observe that a minimal increase in the players' payoff memory (compared to the baseline case with a single round recalled) can promote cooperation considerably. 
Specifically, in all four scenarios with extended memory, we see similar cooperation rates, and they approach the rates observed under perfect memory. 
These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest. 
Already remembering a few interactions, either with the same co-player or across different co-players, may provide players with enough information to adopt reciprocal strategies.

%% RESULTS: Snowdrift Game
{\bf Beyond the donation game.}
So far, we have discussed the donation game, which is the most common
example of a social dilemma. There are several other examples, which include the
snowdrift game~\cite{doebeli:Ecology:2005}, the stag hunt
game~\cite{pacheco:ProcB:2009}, and the public goods
game~\cite{pinheiro:PLoSCompBio:2014}. The pattern we observe is more
general and still holds for other social dilemmas. To demonstrate this, we
consider the snowdrift game. In the snowdrift game, individuals can again
cooperate or defect. However, now if both cooperate, they have the benefit \(b\)
of getting home while sharing the labor \(c\). Thus for mutual cooperation, they
receive \(b - \frac{c}{2}\). However, if one defects, they get \(b\) while
the cooperator gets \(b - c\). By contrast to the donation game, now
the best action depends on the co-player: to defect if the other cooperates, but
to cooperate if the other defects.

@@

Apart from the analytical analysis, we evaluate our results with simulations,
which we do for all cases considered in this manuscript.
What we find is that the qualitative results still hold for the snowdrift game.
Also, in that case, we observe that perfect memory leads to systematically
higher cooperation rates (Supplementary Figures~8, 9). However, now less memory
is needed to approximate the classical limit.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=.75\textwidth]{static/cooperation_rate_over_b_and_beta.pdf}
    \caption{{\bf Evolution of direct reciprocity for different parameter values.} 
    To explore the robustness of our results, we have run simulations for different benefit values (left panels, {\it a}), and for different selection strengths (right panels, {\it b}). 
    In each case, we record the resulting average cooperation rate over the entire simulation (upper panels). 
    In addition we record the individuals' average generosity. 
    Here, we only take into account those residents with $p\! \approx\! 1$ and we compute
    the average of their cooperation probability~$q$. 
    These simulations suggest that perfect payoff memory consistently leads to more cooperation and more generosity. 
    Unless explicitly varied, the parameters of the simulation are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$.
    Each point represents a single simulation run. Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter combination. Each point is a single simulation run.}
    \label{fig:cooperation_rate_over_benefit_and_beta}
\end{figure}
